---
title: "MetricsWeighted"
author: "Michael Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MetricsWeighted}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup}
library(MetricsWeighted)
```

## Introduction

The R package `MetricsWeighted` provides weighted versions of different machine learning metrics, scoring functions and performance measures as well as tools to use it within a `dplyr` chain.

## Installation
From CRAN:
```
install.packages("MetricsWeighted")
```

Latest version from github:
```
library(devtools)
install_github("mayer79/MetricsWeighted")
```

## Illustration

### Metrics, Scoring Functions and Performance Measures

Currently, the following metrics, scoring functions and performance measures are available:

- `accuracy`, `recall`, `precision`, `f1_score`, and `classification_error`: Typical binary performance measures derived from the confusion matrix, see e.g. (https://en.wikipedia.org/wiki/Precision_and_recall). Require binary predictions. Except for classification error, high values are better.

- `AUC` and `gini_coefficient`: Area under the receiver operating curve (ROC) and the closely related Gini coefficient. Actual values must be 0 or 1, while predictions can take any value (only their order is relevant). The higher, the better.

- `deviance_bernoulli` and `logLoss`: Further metrics relevant for binary targets, namely the average unit deviance of the *binary* logistic regression model (0-1 response) and logLoss (half the deviance). As with all deviance measures, smaller values are better.

- `mse`, `mae`, `mape`, and `rmse`: Typical regression metrics (mean-squared error, mean-absolute error, mean absolute percentage error and root-mean-squared error). The lower, the better.

- `deviance_tweedie`: (Unscaled) average unit Tweedie deviance with parameter `tweedie_p`, see [1] and (https://en.wikipedia.org/wiki/Tweedie_distribution) for a reference of the underlying formula.

- `deviance_normal`,  `deviance_gamma`, and `deviance_poisson`: Special cases of Tweedie. `deviance_normal` equals mean-squared error.

They all take four arguments: 

1. `actual`: Actual observed values.

2. `predicted`: Predicted values.

3. `w`: Optional vector with case weights.

4. `...`: Further arguments.

Notable exception is `deviance_tweedie`. It requires the Tweedie power as additional parameter. The value 0 corresponds to the Gaussian/normal deviance, the value 1 to the Poisson deviance and the value 2 to the Gamma deviance, see e.g. [1]. For $0 < p < 1$, the Tweedie distribution is undefined.

For fixed actual and predicted values, the Tweedie deviance is continuous with respect to this parameter.

#### Examples: Regression
```{r}
# The data
y_num <- iris[["Sepal.Length"]]
fit_num <- lm(Sepal.Length ~ ., data = iris)
pred_num <- fit_num$fitted
weights <- seq_len(nrow(iris))

# Performance metrics
mae(y_num, pred_num)  # unweighted
mae(y_num, pred_num, w = rep(1, length(y_num)))  # same
mae(y_num, pred_num, w = weights)  # different
rmse(y_num, pred_num)

# Normal deviance equals Tweedie deviance with parameter 0
deviance_normal(y_num, pred_num)
deviance_tweedie(y_num, pred_num, tweedie_p = 0)
deviance_tweedie(y_num, pred_num, tweedie_p = -0.001)

# Poisson deviance equals Tweedie deviance with parameter 1
deviance_poisson(y_num, pred_num)
deviance_tweedie(y_num, pred_num, tweedie_p = 1)
deviance_tweedie(y_num, pred_num, tweedie_p = 1.01)

# Gamma deviance equals Tweedie deviance with parameter 2
deviance_gamma(y_num, pred_num)
deviance_tweedie(y_num, pred_num, tweedie_p = 2)
deviance_tweedie(y_num, pred_num, tweedie_p = 1.99)
deviance_tweedie(y_num, pred_num, tweedie_p = 2.01)

```

#### Examples: Binary classification
```{r}
# The data
y_cat <- iris[["Species"]] == "setosa"
fit_cat <- glm(y_cat ~ Sepal.Length, data = iris, family = binomial())
pred_cat <- predict(fit_cat, type = "response")

# Performance metrics
AUC(y_cat, pred_cat)  # unweighted
AUC(y_cat, pred_cat, w = weights)  # weighted
logLoss(y_cat, pred_cat)  # Logloss
deviance_bernoulli(y_cat, pred_cat)  # LogLoss * 2

```

### Generalized R-squared

Furthermore, we provide a generalization of R-squared, defined as the proportion of deviance explained, i.e. one minus the ratio of residual deviance and intercept-only deviance, see e.g. [2]. By default, it calculates the ordinary R-squared, i.e. proportion of *normal* deviance (mean-squared error) explained. However, you can specify any different deviance function, e.g. `deviance_tweedie` with paramter 1.5 or the deviance of the *binary* logistic regression (`deviance_bernoulli`).

#### Examples
```{r}
summary(fit_num)$r.squared

# same
r_squared(y_num, pred_num)
r_squared(y_num, pred_num, deviance_function = deviance_tweedie, tweedie_p = 0)
r_squared(y_num, pred_num, deviance_function = deviance_tweedie, tweedie_p = 1.5)

# weighted
r_squared(y_num, pred_num, w = weights)
r_squared(y_num, pred_num, w = weights, deviance_function = deviance_gamma) 
r_squared(y_num, pred_num, w = weights, deviance_function = deviance_tweedie, tweedie_p = 2)
r_squared(y_num, pred_num, deviance_function = deviance_tweedie, tweedie_p = 1.5)

# respect to own deviance formula
myTweedie <- function(actual, predicted, w = NULL, ...) {
  deviance_tweedie(actual, predicted, w, tweedie_p = 1.5, ...)
}
r_squared(y_num, pred_num, deviance_function = myTweedie)

```

### Tidyverse

In order to facilitate the use of these metrics in a `dplyr` chain, you can try out the function `performance`: Starting from a data set with actual and predicted values (and optional case weights), it calculates one or more metrics. The resulting values are returned as a `data.frame`. Stratified performance calculations can e.g. be done by using `do` from `dplyr`.

#### Examples
```{r}
require(dplyr)

# Regression with `Sepal.Length` as response
iris %>% 
  mutate(pred = predict(fit_num, data = .)) %>% 
  performance("Sepal.Length", "pred")

# Same
iris %>% 
  mutate(pred = predict(fit_num, data = .)) %>% 
  performance("Sepal.Length", "pred", metrics = rmse)

# Grouped by Species
iris %>% 
  mutate(pred = predict(fit_num, data = .)) %>% 
  group_by(Species) %>% 
  do(performance(data = ., actual = "Sepal.Length", predicted = "pred"))

# Customized output
iris %>% 
  mutate(pred = predict(fit_num, data = .)) %>% 
  performance("Sepal.Length", "pred", value = "performance",
              metrics = list(`root-mean-squared error` = rmse))

# Multiple measures
iris %>% 
  mutate(pred = predict(fit_num, data = .)) %>% 
  performance("Sepal.Length", "pred",
              metrics = list(rmse = rmse, mae = mae, `R-squared` = r_squared))

# Grouped by Species
iris %>% 
  mutate(pred = predict(fit_num, data = .)) %>% 
  group_by(Species) %>% 
  do(performance(., "Sepal.Length", "pred",
                 metrics = list(rmse = rmse, mae = mae, `R-squared` = r_squared)))

# Passing extra argument (Tweedie p)
iris %>% 
  mutate(pred = predict(fit_num, data = .)) %>% 
  performance("Sepal.Length", "pred",
              metrics = list(`normal deviance` = deviance_normal, 
                             `Tweedie with p = 0` = deviance_tweedie),
              tweedie_p = 0)

```

## References

[1] Jorgensen, B. (1997). The Theory of Dispersion Models. Chapman & Hall/CRC. ISBN 978-0412997112.

[2] Cohen, J. et al. (2002). Applied Multiple Regression/Correlation Analysis for the Behavioral Sciences (3rd ed.). Routledge. ISBN 978-0805822236.
